---
title: "run_analysis"
author: "Alireza Haddadi"
date: "2022-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### This R markdown is about a 5-step procces to make the raw data into Tidy data for the project of **"Getting and Cleaning Data"** course in [Coursera.org](https://www.coursera.org/learn/data-cleaning/home/info).

<br/>

At first we need "dplyr" package, so if the package was not downloaded in your RStudio, make sure to install it, else just run the library function.

```{r message=FALSE, warning=FALSE}
library(dplyr)
```

<br/>

The project gives us the URL for a zip file. The file will be downloaded and saved as "UCI HAR Dataset" folder in your project folder.

```{r}
Url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"

if (!file.exists("UCI HAR Dataset")){
        temp <- tempfile()
        download.file(Url, temp)
        unzip(temp)
        unlink(temp)
}
```

<br/> The whole dataset is separated as the below files and an appropriate column name is set for each file.

```{r}
features <- read.table("UCI HAR Dataset/features.txt", col.names = c("n", "functions"))
activities <- read.table("UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
subject_test <- read.table("UCI HAR Dataset/test/subject_test.txt", col.names = "subject")
x_test <- read.table("UCI HAR Dataset/test/X_test.txt", col.names = features$functions)
y_test <- read.table("UCI HAR Dataset/test/y_test.txt", col.names = "code")
subject_train <- read.table("UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
x_train <- read.table("UCI HAR Dataset/train/X_train.txt", col.names = features$functions)
y_train <- read.table("UCI HAR Dataset/train/y_train.txt", col.names = "code")
```

#### Now it is time to run the "run_analysis.R" in my created repository, and I explained each steps to get the results that project wants from us.

### step 1.

Now we want the whole dataset from each file in step 2. to save as "Merge_data".

```{r}
X <- rbind(x_train, x_test)
Y <- rbind(y_train, y_test)
Subject <- rbind(subject_train, subject_test)
Merged_Data <- cbind(Subject, Y, X)
```

### step 2.

This project only wants the "mean" and "std" functions to be applied to Dataset, so we get the coulmns of which "mean" and "std" are represented.

```{r}
TidyData <- Merged_Data %>% select(subject, code, contains("mean"), contains("std"))
```

### step 3.

Give a specific name for every activity from "activity" file.

```{r}
TidyData$code <- activities[TidyData$code, 2]
```

### step 4.

adjust the column names of "TidyData" in an appropriate way.

```{r}
names(TidyData)[2] = "activity"
names(TidyData)<-gsub("Acc", "Accelerometer", names(TidyData))
names(TidyData)<-gsub("Gyro", "Gyroscope", names(TidyData))
names(TidyData)<-gsub("BodyBody", "Body", names(TidyData))
names(TidyData)<-gsub("Mag", "Magnitude", names(TidyData))
names(TidyData)<-gsub("^t", "Time", names(TidyData))
names(TidyData)<-gsub("^f", "Frequency", names(TidyData))
names(TidyData)<-gsub("tBody", "TimeBody", names(TidyData))
names(TidyData)<-gsub("-mean()", "Mean", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("-std()", "STD", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("-freq()", "Frequency", names(TidyData), ignore.case = TRUE)
names(TidyData)<-gsub("angle", "Angle", names(TidyData))
names(TidyData)<-gsub("gravity", "Gravity", names(TidyData))
```

### step 5.

With the "TidyData" from step 4., "FinalData" is created while we apply mean function to every columns.

```{r, warning = FALSE}
FinalData <- TidyData %>%
        group_by(subject, activity) %>%
        summarise_all(funs(mean))
write.table(FinalData, "FinalData.txt", row.name=FALSE)

str(FinalData)
```
